KidsAIR: Main analysis initial steps

### 2020-10-27 main_analysis_initial.Rmd
analysis upates new RMD for primary models

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

@@

```{r}
# clean up data for table 1 summary
table_1_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  filter(area == "NN") %>% 
  distinct(area, home, .keep_all = TRUE)

@@

```{r}
# clean up data for table 1 summary
table_1_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  filter(area == "NN") %>% 
  distinct(area, home, .keep_all = TRUE)

@@

at_home_pm_summary <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  mutate(new_var = pm_at_home_sampling_period) %>% 
  filter(!is.na(new_var)) %>%
  filter(area == "NN") %>% 
  group_by(treatment) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
at_home_pm_summary
```

```{r}
# Initial summary data for LRTI
incidence_summary <- analysis_data %>% 
  filter(area == "NN") %>% 
  #group_by(treatment) %>% 
  #filter(lrti_incidence < 0.7) %>% 
  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
            sd_inc = sd(lrti_incidence, na.rm = TRUE),
            n_inc = n(),
            min_inc = min(lrti_incidence, na.rm = TRUE),
            median_inc = median(lrti_incidence, na.rm = TRUE),
            max_inc = max(lrti_incidence, na.rm = TRUE))
incidence_summary

count_summary <- analysis_data %>% 
  group_by(treatment) %>% 
  filter(lrti_incidence < 0.7) %>% 
  summarize(mean_events = mean(lrti_events, na.rm = TRUE),
            sd_events = sd(lrti_events, na.rm = TRUE),
            n_events = n(),
            min_events = min(lrti_events, na.rm = TRUE),
            median_events = median(lrti_events, na.rm = TRUE),
            max_events = max(lrti_events, na.rm = TRUE))
count_summary

# amb data summary
# amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti
ambient_summary <- analysis_data %>% 
  filter(area == "NN") %>% 
  mutate(new_var = amb_pm_lrti) %>% 
  group_by(treatment) %>% 
  filter(!is.na(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  summarize(mean_amb = mean(new_var, na.rm = TRUE),
            sd_amb = sd(new_var, na.rm = TRUE),
            n_amb = n(),
            min_amb = min(new_var, na.rm = TRUE),
            median_amb = median(new_var, na.rm = TRUE),
            max_amb = max(new_var, na.rm = TRUE))
ambient_summary
```

```{r}
# Histogram for distribution of data
events_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events)) +
    theme_light() +
    facet_wrap(~treatment)
events_plot

@@

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events_di)) +
    facet_wrap(~treatment)
events_di_plot
```

```{r}
# Summarize initial models, check overdispersion and other diagnostic criteria
library(MASS)

@@

library(pscl)
odTest(nb_results) # test indicates poisson model is overdispersed
# According to these results, data is overdispersed
# Variance is much higher than the mean for LRTI incidence
# Assumption of Poisson is variance = mean, so recommend using negative binomial model

@@

library(faraway)
halfnorm(residuals(negbin_results))

```

>>>>

```{r}
# Load analysis dataset

@@

```{r}
# Summary analysis for categorical variables

# clean up data for table 1 summary
table_1_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  filter(area == "NN") %>% # Change area filter as necessary
  distinct(area, home, .keep_all = TRUE)

@@

```{r}
# Summary analysis for numeric variables

# clean up data for table 1 summary
table_1_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  filter(area == "NN") %>% # Change area filter as necessary
  distinct(area, home, .keep_all = TRUE)

@@

# At-home PM summary - different than function above due to distinct obs by child instead of home
at_home_pm_summary <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  mutate(new_var = pm_at_home_sampling_period) %>% 
  filter(!is.na(new_var)) %>%
  filter(area == "NN") %>% # Change area filter as necessary 
  group_by(treatment) %>% # Change treatment grouping as necessary 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
at_home_pm_summary
```

```{r}
# Summary for outcome events and ambient data

# Initial summary data for LRTI
incidence_summary <- analysis_data %>% 
  filter(area == "NN") %>% # Change area filter as necessary  
  #group_by(treatment) %>% # Change treatment grouping as necessary  
  #filter(lrti_incidence < 0.7) %>% 
  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
            sd_inc = sd(lrti_incidence, na.rm = TRUE),
            n_inc = n(),
            min_inc = min(lrti_incidence, na.rm = TRUE),
            median_inc = median(lrti_incidence, na.rm = TRUE),
            max_inc = max(lrti_incidence, na.rm = TRUE))
incidence_summary

count_summary <- analysis_data %>% 
  group_by(treatment) %>% # Change treatment grouping as necessary 
  #filter(lrti_incidence < 0.7) %>% 
  summarize(mean_events = mean(lrti_events, na.rm = TRUE),
            sd_events = sd(lrti_events, na.rm = TRUE),
            n_events = n(),
            min_events = min(lrti_events, na.rm = TRUE),
            median_events = median(lrti_events, na.rm = TRUE),
            max_events = max(lrti_events, na.rm = TRUE))
count_summary

# amb data summary
# amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti
ambient_summary <- analysis_data %>% 
  filter(area == "NN") %>% # Change area filter as necessary  
  mutate(new_var = amb_pm_lrti) %>% 
  group_by(treatment) %>% # Change treatment grouping as necessary 
  filter(!is.na(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% # convert degrees F to C 
  summarize(mean_amb = mean(new_var, na.rm = TRUE),
            sd_amb = sd(new_var, na.rm = TRUE),
            n_amb = n(),
            min_amb = min(new_var, na.rm = TRUE),
            median_amb = median(new_var, na.rm = TRUE),
            max_amb = max(new_var, na.rm = TRUE))
ambient_summary
```

```{r}
# Histograms for distribution of outcome data

events_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events)) +
    theme_light() +
    facet_wrap(~treatment)
events_plot

@@

events_plot_pm <- analysis_data %>% 
  mutate(pm_cat = cut(pm_mean_sampling_period,
                      breaks = c(0, 9.85, 18.06, 34.88, 286),
                      labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
  filter(!is.na(pm_cat)) %>% 
  ggplot() +
    geom_col(aes(pm_cat, lrti_events)) +
    theme_light() +
    facet_wrap(~treatment)
events_plot_pm

scatter_plot_pm <- analysis_data %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  ggplot() +
    geom_point(aes(lrti_events, pm_mean_sampling_period)) +
    theme_light() +
    facet_wrap(~treatment)
scatter_plot_pm

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events_di)) +
    facet_wrap(~treatment)
events_di_plot

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_boxplot(aes(lrti_events_di, pm_mean_sampling_period, group = lrti_events_di)) +
    facet_wrap(~treatment)
events_di_plot
```

```{r}
# Summarize initial models, check overdispersion and other diagnostic criteria

library(MASS)

@@

library(pscl)
odTest(nb_results) # test indicates poisson model is overdispersed

# According to these results, data is overdispersed
# Variance is much higher than the mean for LRTI incidence
# Assumption of Poisson is variance = mean, so recommend using negative binomial model

@@

library(faraway)
halfnorm(residuals(negbin_results))
```
